Search-based Synthesis of Network Intrusion Detection Rules

### 2019-09-23 paper/paper.tex
pass on intro

\sloppy \nids{} are typically categorized in two
groups~\cite{kumar2007survey}: 1)~rule-based and 2)~anomaly-based. A
rule-based \nids{} (\aka\ signature-based NIDS) checks if the network
traffic matches a fixed set of
rules. Figure~\ref{fig:synflood-example} shows an example rule of
Suricata~\cite{suricata}, a popular open-source \nids{}. This rule
prescribes a method to capture SYN flood attacks to a
server~\cite{Douligeris:2004:DAD:987153.987158} that can result in
denial of service. Rule-based \nids{} focuses on known attacks whereas
anomaly-based \nids{} focuses on unanticipated attacks that could be
observed through suspicious network
traffic~\cite{kumar2007survey,Mitchell:2014:SID:2597757.2542049,cordy-etal-issta19}.

Rule-based and Anomaly-based NIDS are complementary. This paper
proposes a technique, dubbed \tname{}, to automatically synthesize
rules from positive and negative examples (\ie{}, benign and malicious
traffic). The main scenario of application is one where an
anomaly-based NIDS identifies abnormal traffic and signals that
traffic to \tname{} to craft rules that can be used by rule-based
NIDS. The core motivation is that attackers are productive in creating
new ways to circumvent existing protections and manual creation of
rules is tedious and time-consuming.\Mar{Lucas/Guilherme: (IMP) do we
  have data to support this?} For those reasons, synthesizing rules to
circumvent those issues is an important problem.

@@

@@

\begin{figure}[t]
  \lstinputlisting[language=C,numbers=none]{synflood.suricata}
  \caption{Suricata rule for SYN Flood Attacks.\Mar{colocar
      propriedades em negrito}}
  \label{fig:synflood-example}
\end{figure}

>>>>

\sloppy \nids{} are typically categorized in two
groups~\cite{kumar2007survey}: 1)~rule-based and 2)~anomaly-based. A
rule-based \nids{} (\aka\ signature-based NIDS) checks if the network
traffic matches a fixed set of
rules. Figure~\ref{fig:synflood-example} shows an example rule of
Suricata~\cite{suricata}, a popular open-source \nids{}. This rule
prescribes a method to capture a denial-of-service attack to a
server~\cite{Douligeris:2004:DAD:987153.987158}. Rule-based \nids{} focuses on known attacks whereas
anomaly-based \nids{} focuses on unanticipated attacks that could be
observed through suspicious network
traffic~\cite{kumar2007survey,Mitchell:2014:SID:2597757.2542049,cordy-etal-issta19}.

These intrusion detection mechanisms are complementary. Rule-based
NIDS are very popular in industry (it is clear to system
administrators what rules will do), but attacker are creating in
evading the current set of rules, and writing these rules is tedious
and error-prone. Anomaly-based NIDS are not restricted to a fixed set
of rules; they learn common behaviors and dynamically detect uncommon
behavior. The often use some machine learning algorithm to model
sequence of events~\cite{7579764}. Unfortunately, anomaly-based NIDS
can raise false alarms as behavior models are fundamentally
imprecise. To sum up, both approaches have positive and negative
aspects.

%% This paper focuses on the
%% problem of discovering new rules by using anomaly-based NIDS.

This paper proposes a technique, dubbed \tname{}, to automatically
synthesize rules from positive and negative examples (\ie{}, benign
and malicious traffic). The main scenario of application is one where
an anomaly-based NIDS identifies abnormal traffic and signals that
traffic to \tname{} to craft rules that can be used by rule-based
NIDS. Those rules can be distributed to rule databases. The core
motivation is that 1) attackers are productive in creating new ways to
circumvent existing protections and 2) manual creation of rules is
tedious and time-consuming.  Note that anomaly-based NIDS are evolving
pretty quick with the advances in machine learning, but rule-based
NIDS are still extremely popular. Our goal is to harden rule-based
NIDS by automatically creating rules from suspicious behavior detected
by anomaly-based NIDS.

@@

\begin{figure}[t]
  \lstinputlisting[language=C,numbers=none,keywords={flags,threshold}]{synflood.suricata}
  \caption{Suricata rule for SYN Flood Attacks.}
  \label{fig:synflood-example}
\end{figure}

@@
